%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Maxime Ulysse Garcia, max@ithake.eu, 2013 %%%
%%%                                           %%%
%%%    Questions and suggestions WELCOMED!    %%%
%%%                                           %%%
%%% This thesis template largely derives from %%%
%%%      Charles Chapple, Robert Castelo      %%%
%%%    Sergio Mendoza and Sergi Castellano    %%%
%%%       Under GNU/GPL copyleft license      %%%
%%%                                           %%%
%%%   FEEL FREE TO USE IT AND IMPROVE IT!!!   %%%
%%%                                           %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Preamble: LaTeX parameters and extensions %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%               Document type               %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% While working to see output errors
%\documentclass[draft,pdftex,oneside]{book}

% To produce final printable version
\documentclass[pdftex,oneside]{book}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%              Tweaking errors              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% To remove "No room for a new \dimen" when using tikz
\RequirePackage{etex}
% For fancyhdr
\setlength{\headheight}{13pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%             Our own style file            %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Style file for this template myThesis.sty
\RequirePackage{myThesis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                   Colors                  %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage[usenames,
	dvipsnames,
	svgnames,
	table]{xcolor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                Paper layout               %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 

\RequirePackage[paperwidth=170mm,
	paperheight=240mm,
	hmargin={2cm,1cm},
	vmargin={2cm,1.75cm},
	headheight=14pt]{geometry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                  Headings                 %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%For memoir class
\let\footruleskip\undefined

% To produce nice headings
\RequirePackage{fancyhdr}
\pagestyle{fancy}

% Big first letter.
\RequirePackage{lettrine}
\newcommand{\mylettrine}[1]{\lettrine[lines=3,lhang=.1]{#1}}

% Fancy looking boxes
\RequirePackage{tcolorbox}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%            Marks and draftmarks           %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\myversion}{V 0.7}

\renewcommand{\chaptermark}[1]{\markboth{\thechapter. #1}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}

\lfoot{}
\rfoot{}

% Normal marks
%\lhead[\fancyplain{}{\sffamily\thepage}]{\fancyplain{}{\sffamily\rightmark}}
%\rhead[\fancyplain{}{\sffamily\leftmark}]{\fancyplain{}{\sffamily\thepage}}
%\cfoot{}

%Draftmarks
\lhead[\fancyplain{}{\sffamily\thepage}]{\fancyplain{}{\sffamily\textcolor{blue}{DRAFT} \rightmark}}
\rhead[\fancyplain{}{\sffamily\leftmark}]{\fancyplain{}{\sffamily\thepage}}
\cfoot{\textcolor{blue}{\myversion}}

%%% Line Spacing %%%
\RequirePackage{setspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                PDF commands               %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% To produce final version
\RequirePackage{pdfpages}

% Used in each \%includepdf call unless overwritten locally
\includepdfset{frame=false,pagecommand={\thispagestyle{fancy}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                  Figures                  %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Path to figures
\graphicspath{{figures//}}
% To include graphics
\RequirePackage{graphicx}
% To wrap text
\RequirePackage{wrapfig}
% to draw Figures
\RequirePackage{tikz}
% to rotate figures and tables
\RequirePackage{rotating}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                 Equations                 %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage{mathtools}

% Redefine the square root:
\let\oldsqrt\sqrt
\def\sqrt{\mathpalette\DHLhksqrt}
\def\DHLhksqrt#1#2{
	\setbox0=\hbox{$#1\oldsqrt{#2\,}$}\dimen0=\ht0
	\advance\dimen0-0.2\ht0
	\setbox2=\hbox{\vrule height\ht0 depth -\dimen0}
	{\box0\lower0.4pt\box2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                Big numbers                %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage{numprint}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                   Lists                   %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Provides a new environnement to do a list
\newenvironment{mylist}[1]{
	\begin{list}{}{
		\settowidth{\labelwidth}{#1:}
		\setlength{\labelsep}{0.5cm}
		\setlength{\leftmargin}{\labelwidth}
		\addtolength{\leftmargin}{\labelsep}
		\setlength{\rightmargin}{0pt}
		\setlength{\parsep}{0.5ex plus 0.2ex minus 0.1ex}
		\setlength{\itemsep}{0 ex plus 0.2ex}
		\renewcommand{\makelabel}[1]{\textbf{##1}\hfil}
		}
	}
{\end{list}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                  Acronyms                 %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage{suffix}
\RequirePackage[printonlyused]{acronym}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                  Verbatim                 %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage{moreverb}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                 References                %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage[round,comma,numbers,super,sort&compress]{natbib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                   Index                   %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage{makeidx}
\makeindex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                   Fonts                   %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage{palatino}
\RequirePackage{amsfonts}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%               URL inclusion               %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage[pdftex,
	colorlinks=true,
	linkcolor=black,
	menucolor=red,
	pdfhighlight=/P,
	citecolor=MidnightBlue,
	urlcolor=BrickRed,
	linktocpage=false]{hyperref}
%Enter URLs like this: \url{http://ithake.eu/}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%               Defining terms              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\myauthor{Maxime Ulysse Garcia}

\def\mytitlefr{Découverte de biomarqueurs prédictifs en cancer du sein par Intégration Transcrip\-tome-Interactome}

\def\mytitleen{Biomarkers discovery in breast cancer by Interactome-Transcriptome Integration}

\def\myabstractfr{L'arrivée des technologies à haut-débit pour mesurer l'expression des gènes a permis l'utilisation de signatures génomiques pour prédire des conditions cliniques ou la survie du patient.
Cependant de telles signatures ont des limitations, comme la dépendance au jeu de données d'entrainement et le manque de généralisation.
Nous proposons un nouvel algorithme, \ac{ITIfr} (\citet{Garcia2012}) pour extraire une signature generalisable prédisant la rechute métastatique dans le cancer du sein par superimposition d'un très large jeu de données d'interaction protèine-protèine sur de multiples jeux de données d'expression des gènes.
Cette méthode ré-implemente l'algorithme \citet{Chuang2007}, avec la capacité supplémentaire d'extraire une signature génomique à partir de plusieurs jeux de donnés d'expression des gènes simultanément.
Deux études ont été faites sur cinq jeux de donnés issus de puces à ADN en cancer du sein.
Pour chaque étude, deux signatures séparées ont été établies sur les patients sur-exprimant les récepteur aux {\oe}strogènes (ER+) and les patients n'exprimant pas les récepteur aux {\oe}strogènes (ER-). 
Nous aovns trouvé deux signatures composées de 6 et 139 sous-réseaux reliées respectivement à la rechute métastatique à 5 ans dans le cancer du sein ER+ et ER-.
Ces signatures ont été confrontées aux signatures \citet{Wang2005} et \citet{vantveer2002} contre des jeux de données indépendants.
Contre le jeu de données \citet{Desmedt2008}, dans les ER+/ER- la précision est respectivement de 0.74/0.54, 0.41/0.44 et 0.60/0.38 pour les signatures ITI, Van’t Veer et Wang.
Contre le jeu de données \citet{vandevijver2002}, dans les ER+/ER- la précision est respectivement de 0.52/0.53, 0.62/0.53 et 0.63/0.56 pour les signatures ITI, Van’t Veer et Wang.
Nous avons trouvés des sous-réseaux formant des complexes précédement relié à des fonctions biologiques impliquées dans la nétastase et le cancer du sein.
Plusieurs gènes directeurs ont été détectés, dont \acs{CDK1}, \acs{NCK1} et \acs{PDGFB}, certains n'étant pas déjà relié à la rechute métastatique dans le cancer du sein.}

\def\myabstracten{High-throughput gene-expression profiling technologies yeild genomic signatures to predict clinical condition or patient outcome.  
However, such signatures have limitations, such as dependency on training set, and lack of generalization.  
We propose a novel algorithm, \ac{ITIen} (\citet{Garcia2012}) to extract a generalizable signature predicting breast cancer relapse by superimposition of a large-scale protein-protein interaction data over several gene-expression data sets.  
This method re-implements the \citet{Chuang2007} algorithm, with the added capability to extract a genomic signature from several gene-expression data sets simultaneously.  
Two studies were made with five breast cancer DNA microarray data sets.  
For each studies, two separate signatures were established on patients overexpressing Estrogen Receptor (ER+) and patients non-expressing Estrogen Receptor (ER-).  
We found two sets of 6 and 139 subnetworks signatures linked respectively to 5 years relapse free survival in breast cancer ER+ and ER-.  
These were confronted to \citet{Wang2005} and \citet{vantveer2002} signatures against independent datasets.  
Against \citet{Desmedt2008} Dataset , in ER+/ER- the accuracy is respectively of 0.74/0.54, 0.41/0.44 and 0.60/0.38 for ITI, Van’t Veer and Wang signatures.  
Against \citet{vandevijver2002} Dataset, in ER+/ER- the accuracy is respectively of 0.52/0.53, 0.62/0.53 and 0.63/0.56 for ITI, Van’t Veer and Wang signatures.  
We found that subnetworks formed complexes functionally linked to biological functions related to metastasis and breast cancer.  
Several drivers genes were detected, including \acs{CDK1}, \acs{NCK1} and \acs{PDGFB}, some not previously linked to breast cancer relapse.}

\def\mykeywordsfr{Transcriptome; Interactome; Intégration de données; Réseaux de gène; classification; SVM; Signature; Biomarqueurs; Cancer; Cancer du Sein}

\def\mykeywordsen{Transcriptome; Interactome; Data Integration; Gene Networks; classification; SVM; Signature; Biomarkers; Cancer; Breast Cancer}

% To get subscript easily
\newcommand{\textunderscript}[1]{$_{\text{#1}}$}

% To get e and er in superscript easily
\newcommand{\eme}{\textsuperscript{e} }
\newcommand{\er}{\textsuperscript{er} }

% To get CO2 and O2 easily
\newcommand{\COO}{CO\textunderscript{2} }
\newcommand{\OO}{O\textunderscript{2} }

% shorther command for boldface
\newcommand{\bb}[1]{\textbf{#1}}

%%% to add an empty page, output pending figures and resume in an odd-numbered one %%%

\newcommand{\clearemptypage}{\newpage\thispagestyle{empty}\mbox{}}

\newcommand{\clearemptydoublepage}{
	\clearemptypage
	\if@twoside
		\ifodd
			\clearemptypage
		\fi
	\fi}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                 PDF infos                 %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypersetup{
	pdftitle		= {\mytitlefr},
	pdfsubject		= {Thèse de doctorat en Bioinformatique et Génomique à Aix-Marseille Université},
	pdfkeywords		= {\mykeywordsfr},
	pdfauthor		= {\myauthor},
	pdfcreator		= {\LaTeX},
	pdfproducer		= {pdfeTeX-0.\the\pdftexversion\pdftexrevision}
	pdfpagemode		= FullScreen,
	pdffitwindow	= true}

\AtBeginDocument{\pdfpagewidth=176mm\pdfpageheight=250mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                 Paragraphs                %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setlength{\parskip}{1ex plus0.2ex minus0.2ex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                 Footnotes                 %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Vertical spacing between footnotes
\setlength{\footnotesep}{2ex}

% Rule height, width and space
\renewcommand{\footnoterule}{\noindent\rule{5cm}{.1ex}\vspace{1.5ex}}

% That's all folks!